---
title: "normal cel subset"
output: html_notebook
---

```{r}

```


```{r}
normal_final_srt <- readRDS("~/CASCADEpaper/paper/Fig6/normal_final_srt.Rds")

normal_final_srt$cell_id <- factor(normal_final_srt$cell_id, levels = c("T cell","B cell","Plasma cell",  "Naive cell", "Macrophage", "Erythroblast", "Epithelial cell", "Adipocyte","Chondrocyte", "Smooth muscle", "Fibroblast", "Endothelial cell", "Hepatocyte", "Neuron"))
DotPlot(normal_final_srt, features = unique(all_marker), group.by = "cell_id")+ theme(axis.text.x = element_text(angle = 90, hjust = 1))
DimPlot(normal_final_srt, label = T)

```
```{r harmony on RNA }
harmony_rna <- function(endo, res = 0.5){

DefaultAssay(endo) <- "RNA"
endo <- NormalizeData(endo)
endo <- ScaleData(endo)

endo <- FindVariableFeatures(endo, selection.method = "vst", nfeatures = 2000)
endo <- RunPCA(endo, features = VariableFeatures(object = endo))
endo <- RunHarmony(endo, "patient", assay.use = "RNA")
endo <- endo %>% 
  RunUMAP(reduction = "harmony", dims = 1:20) %>% 
  FindNeighbors(reduction = "harmony", dims = 1:20) %>% 
  FindClusters(resolution = res) %>% 
  identity()

p <- DimPlot(endo, label = T) + DimPlot(endo, label = T, group.by = "patient")
print(p)
return(endo)

}
```

```{r}

library(gridExtra)
library(DOSE)
library(msigdbr)
library(clusterProfiler)
library(enrichplot)
library(ggnewscale)
library(dorothea)
library(dplyr)
library(tibble)
library(data.table)
Hall <- msigdbr(species = "Homo sapiens", category = "H") %>%
  dplyr::select(gs_name, gene_symbol)
KEGG <- msigdbr(species = "Homo sapiens", category = "C2", subcategory = "KEGG")%>%
  dplyr::select(gs_name, gene_symbol)
Reactome <- msigdbr(species = "Homo sapiens", category = "C2", subcategory = "REACTOME")%>%
  dplyr::select(gs_name, gene_symbol)

C5 <- msigdbr(species = "Homo sapiens", category = "C5") %>%
  dplyr::select(gs_name, gene_symbol)
C6 <- msigdbr(species = "Homo sapiens", category = "C6") %>%
  dplyr::select(gs_name, gene_symbol)
C7 <- msigdbr(species = "Homo sapiens", category = "C7") %>%
  dplyr::select(gs_name, gene_symbol)
all <- rbind(Hall, KEGG, Reactome,C5,C6)

C8 <- msigdbr(species = "Homo sapiens", category = "C8") %>%
  dplyr::select(gs_name, gene_symbol)


enrichment_plots <- function(geneList, geneSets, upregulatetitle, downregulatetitle){
  em2 <- GSEA(geneList, TERM2GENE = geneSets)
  em2_up <- em2
  em2_up@result <- em2_up@result[em2_up@result$NES > 0,]
  if(nrow(em2_up@result)>0){
    d <- dotplot(em2_up, showCategory=30) + ggtitle(upregulatetitle)
    print(d)
  }else { print("No upregulated result")}
  
  em2_down <- em2
  em2_down@result <- em2_down@result[em2_down@result$NES < 0,]
  if(nrow(em2_down@result)>0){
    d <- dotplot(em2_down, showCategory=30) + ggtitle(downregulatetitle)
    print(d)}else{ print("No downregulated result")}
  return(em2)
}

generate_genelist <- function(markerdf){
  temp <- setNames(markerdf$avg_log2FC, rownames(markerdf))
  temp <- sort(temp, decreasing = T)
  return(temp)
}
```


## endothelial cell
```{r message=FALSE, warning=FALSE}
endo <- subset(normal_final_srt, cell_id == "Endothelial cell")
endo <- harmony_rna(endo)
```
n = 2053
```{r}
FeaturePlot(endo, c("PECAM1","CD34"))
FeaturePlot(endo, c("ACKR1", "SELE", "SELP", "ICAM1", "VCAM1", "HLA-DRA"))
FeaturePlot(endo, c("DLL4", "RGS5", "ESM1"))
FeaturePlot(endo, c("DLL4", "CXCL12", "VEGFC"))
FeaturePlot(endo, c("VWF", "GJA5", "DSRED", "NRG1", "EMCN"))
FeaturePlot(endo, c("GJA5", "FBLN5", "LTBP4")) # aterial
FeaturePlot(endo, c("ACKR1", "SELP")) # venous
FeaturePlot(endo, c("CA4", "FCN3", "EDN1")) # capilary
FeaturePlot(endo, c("CXCR4", "ESM1", "ANGPT2")) # tip like
```

```{r  enrichment , fig.height=12, fig.width=8}
endo_marker <- FindAllMarkers(endo)
markerlist <- lapply(split(endo_marker, endo_marker$cluster), generate_genelist)
endo_en <- lapply(names(markerlist), function(cluster) {
  x <-markerlist[[cluster]]
  enrichment_plots(x, C8, paste(cluster, "UP"), paste(cluster, "DOWN"))
  })

```
```{r}
names(endo_en) <- names(markerlist)
endo_en <- endo_en[sapply(endo_en, nrow)>0]
endo_en <- rbindlist(lapply(endo_en, function(x) x@result), idcol = "cluster")
table(endo$site, endo$seurat_clusters)
DimPlot(endo, group.by = "site")
```


cluster 0: vascular endo
cluster 1: vascular endo / aterial cells
cluster 2: mixture of lymphatic /vascular endo
cluster 3: lymphatic endo 
cluster 4: Liver sinusoidal endothelial cells (LSECs)
cluster 5: vein endo
cluster 6: fibroblast/chondrocyte/SM
cluster 7: hepatocytes???


### clean the endothelial cells
```{r}
hepacell <- WhichCells(endo, idents = "7")
fibrocell <- WhichCells(endo, idents = "6")

endo2 <- subset(endo, idents = 0:5)
endo2 <- harmony_rna(endo2)
DimPlot(endo2, label = T) + DimPlot(endo2, label = T, group.by = "patient")
```

```{r}
FeaturePlot(endo2, c("PECAM1","CD34"))
FeaturePlot(endo2, c("ACKR1", "SELE", "SELP", "ICAM1", "VCAM1", "HLA-DRA"))
FeaturePlot(endo2, c("DLL4", "RGS5", "ESM1"))
FeaturePlot(endo2, c("DLL4", "CXCL12", "VEGFC"))
FeaturePlot(endo2, c("VWF", "GJA5", "DSRED", "NRG1", "EMCN"))
FeaturePlot(endo2, c("GJA5", "FBLN5", "LTBP4")) # aterial
FeaturePlot(endo2, c("ACKR1", "SELP")) # venous
FeaturePlot(endo2, c("CA4", "FCN3", "EDN1")) # capilary
FeaturePlot(endo2, c("CXCR4", "ESM1", "ANGPT2")) # tip like
```
```{r fig.height=12, fig.width=8}
endo2_marker <- FindAllMarkers(endo2)
markerlist <- lapply(split(endo2_marker, endo2_marker$cluster), generate_genelist)
endo2_en <- lapply(names(markerlist), function(cluster) {
  x <-markerlist[[cluster]]
  enrichment_plots(x, C8, paste(cluster, "UP"), paste(cluster, "DOWN"))
  })
names(endo2_en) <- names(markerlist)
endo2_en <- endo2_en[sapply(endo2_en, nrow)>0]
endo2_en <- rbindlist(lapply(endo2_en, function(x) x@result), idcol = "cluster")

```
```{r}
table(endo2$site, endo2$seurat_clusters)
DimPlot(endo2, group.by = "site")
DimPlot(endo2)
FeaturePlot(endo2, Fibroblast_feature)
FeaturePlot(endo2, smooth_muscle)
FeaturePlot(endo2,Epithelial_cell)
FeaturePlot(endo2, focused_B_cell_feature)
```
c0 : endothelial tip cells
c1, c2 : vascular endothelial cell 

c4: lymphatic endo 
c3. c5: LSECS 
c6: vein
c7: artery

c8: Epithelial cell?



```{r}
fibrocell <- c(fibrocell, WhichCells(endo2, idents = 9))
immunecell <- WhichCells(endo2, idents = 10)
endo3 <- subset(endo2, idents = 0:7)
endo3 <- harmony_rna(endo3)
```
```{r fig.height=10, fig.width=6}
endo3_marker <- FindAllMarkers(endo3)
markerlist <- lapply(split(endo3_marker, endo3_marker$cluster), generate_genelist)
endo3_en <- lapply(names(markerlist), function(cluster) {
  x <-markerlist[[cluster]]
  enrichment_plots(x, C8, paste(cluster, "UP"), paste(cluster, "DOWN"))
  })
names(endo3_en) <- names(markerlist)
endo3_en <- endo3_en[sapply(endo3_en, nrow)>0]
endo3_en <- rbindlist(lapply(endo3_en, function(x) x@result), idcol = "cluster")
table(endo3$site, endo3$seurat_clusters)
```
```{r}
FeaturePlot(endo3, Fibroblast_feature)
FeaturePlot(endo3, focused_B_cell_feature)
FeaturePlot(endo3, c("PECAM1","CD34"))
FeaturePlot(endo3, c("ACKR1", "SELE", "SELP", "ICAM1", "VCAM1", "HLA-DRA"))
FeaturePlot(endo3, c("DLL4", "RGS5", "ESM1"))
FeaturePlot(endo3, c("DLL4", "CXCL12", "VEGFC"))
FeaturePlot(endo3, c("VWF", "GJA5", "DSRED", "NRG1", "EMCN"))
FeaturePlot(endo3, c("GJA5", "FBLN5", "LTBP4")) # aterial
FeaturePlot(endo3, c("ACKR1", "SELP")) # venous
FeaturePlot(endo3, c("CA4", "FCN3", "EDN1")) # capilary
FeaturePlot(endo3, c("CXCR4", "ESM1", "ANGPT2")) # tip like
```

### final annotation
```{r}
endo3$cell_anno <- NA
endo3$cell_anno[endo3$seurat_clusters == "0"] <- "Endothelial tip cell"
endo3$cell_anno[endo3$seurat_clusters %in% c("1", "2", "3")] <- "Endothelial cell"
endo3$cell_anno[endo3$seurat_clusters== "6"] <- "Vein endothelial cell"
endo3$cell_anno[endo3$seurat_clusters== "7"] <- "Arterial endothelial cell"
endo3$cell_anno[endo3$seurat_clusters== "4"] <- "Lymphatic endothelial cell"
endo3$cell_anno[endo3$seurat_clusters== "5"] <- "Liver sinusoidal endothelial cell"
Idents(endo3) <- endo3$cell_anno
saveRDS(endo3, "Endothelial.Rds")
DimPlot(endo3)
DimPlot(endo3, group.by = "site")
```
```{r}
df <- endo3@meta.data %>% group_by(site, cell_anno) %>% summarise(n = n()) %>% mutate (freq = n/sum(n))

ggplot(df, aes(x = site, y = freq, fill = cell_anno)) + geom_bar(stat = "identity") + 
  theme(axis.text.x = element_text(hjust = 1, vjust = 0.5, angle = 90)) + 
  scale_fill_manual(values = brewer.dark2(6))
```


## fibroblast , chondrocyte, smooth muscle, adipocyte
```{r message=FALSE, warning=FALSE}
fibro <- subset(normal_final_srt, cell_id %in% c("Fibroblast","Chondrocyte", "Smooth muscle", "Adipocyte"))
fibro <- harmony_rna(fibro, res = 1)
```

```{r}
DimPlot(fibro, group.by = "cell_id")
FeaturePlot(fibro, Fibroblast_feature)
FeaturePlot(fibro, smooth_muscle)
FeaturePlot(fibro, Chondrocyte_feature)
FeaturePlot(fibro, Adipocyte_feature)

FeaturePlot(fibro, c("POSTIN", "CXCL14", "COL3A1"))+ plot_annotation(title = "Active")
FeaturePlot(fibro, c("S100A4", "CFD", "DPT"))+ plot_annotation(title = "Resting")
FeaturePlot(fibro, "CTHRC1") + plot_annotation(title = "CAFs")
FeaturePlot(fibro, c("FAP", "COL1A1", "POSTN" )) + plot_annotation(title = "inflam") # also CAFs

```
c0 CAF
c1, c5 = senescent fibro


```{r fig.height=12, fig.width=8}
fibro_marker <- FindAllMarkers(fibro)
markerlist <- lapply(split(fibro_marker, fibro_marker$cluster), generate_genelist)
fibro_en <- lapply(names(markerlist), function(cluster) {
  x <-markerlist[[cluster]]
  enrichment_plots(x, C8, paste(cluster, "UP"), paste(cluster, "DOWN"))
  })
names(fibro_en) <- names(markerlist)
fibro_en <- fibro_en[sapply(fibro_en, nrow)>0]
fibro_en <- rbindlist(lapply(fibro_en, function(x) x@result), idcol = "cluster")

```
```{r}
fibro_marker %>% group_by(cluster) %>% filter(avg_log2FC > 1.5 ) %>% View(.)
fibro_marker %>% group_by(cluster) %>% filter(avg_log2FC > 1.5 ) %>% pull(gene) %>% unique(.) -> DEgenes
DoHeatmap(fibro, DEgenes)
```

```{r}
FeaturePlot(fibro, c("ACTA2", "MYH11", "KCNJ8", "CNN1"), ncol = 3)  # smooth muscle
FeaturePlot(fibro, c("IBSP", "SP7"))
FeaturePlot(fibro, c("IGF1", "C3", "ABCA9")) # skeletal muscle
```

```{r}
immunecell <- c(immunecell, WhichCells(fibro, idents = "11"))
neurocell <- WhichCells(fibro, idents = "12")
fibro2 <- subset(fibro, idents =  c(0:10, 13))
fibro2 <- harmony_rna(fibro2)
```

```{r}
FeaturePlot(fibro2, c("ACTA2", "MYH11", "KCNJ8", "CNN1"), ncol = 3)  # smooth muscle
FeaturePlot(fibro2, c("IBSP", "SP7")) #chondrocyte
FeaturePlot(fibro2, c("IGF1", "C3", "ABCA9")) # skeletal muscle
FeaturePlot(fibro2, c("S100A4", "CFD", "DPT"))+ plot_annotation(title = "Resting")
FeaturePlot(fibro2, c("FAP", "COL1A1", "POSTN","CTHRC1")) + plot_annotation(title = "inflam") # also CAFs
FeaturePlot(fibro2, Adipocyte_feature)
FeaturePlot(fibro2, Fibroblast_feature)
FeaturePlot(fibro2, Neuron_feature)
```

```{r fig.height=12, fig.width=8}
fibro2_marker <- FindAllMarkers(fibro2)
markerlist <- lapply(split(fibro2_marker, fibro2_marker$cluster), generate_genelist)
fibro2_en <- lapply(names(markerlist), function(cluster) {
  x <-markerlist[[cluster]]
  enrichment_plots(x, C8, paste(cluster, "UP"), paste(cluster, "DOWN"))
  })
names(fibro2_en) <- names(markerlist)
fibro2_en <- fibro2_en[sapply(fibro2_en, nrow)>0]
fibro2_en <- rbindlist(lapply(fibro2_en, function(x) x@result), idcol = "cluster")

```

```{r}
#fibro2_marker %>% filter(cluster == "7") %>% View()
fibro2$cell_anno <- "Fibroblast"
fibro2$cell_anno[fibro2$seurat_clusters == "1"] <- "Senescent fibroblast"
fibro2$cell_anno[fibro2$seurat_clusters %in% c("0")] <- "CAFs"
fibro2$cell_anno[fibro2$seurat_clusters== "9"] <- "Skeletal muscle"
fibro2$cell_anno[fibro2$seurat_clusters== "4"] <- "Chondrocyte"
fibro2$cell_anno[fibro2$seurat_clusters== "6"] <- "Smooth muscle"
fibro2$cell_anno[fibro2$seurat_clusters== "3"] <- "Adipocyte"
Idents(fibro2) <- fibro2$cell_anno
saveRDS(fibro2, "Fibroblast.Rds")
DimPlot(fibro2)
DimPlot(fibro2, group.by = "site")
```
```{r}
df <- fibro2@meta.data %>% group_by(site, cell_anno) %>% summarise(n = n()) %>% mutate (freq = n/sum(n))

ggplot(df, aes(x = site, y = freq, fill = cell_anno)) + geom_bar(stat = "identity") + 
  theme(axis.text.x = element_text(hjust = 1, vjust = 0.5, angle = 90)) + 
  scale_fill_manual(values = brewer.dark2(7))

df <- fibro2@meta.data %>% group_by(sample, cell_anno) %>% summarise(n = n()) %>% mutate (freq = n/sum(n))
df$patient <- substr(df$sample, 1, 6 )
ggplot(df, aes(x = sample, y = freq, fill = cell_anno)) + geom_bar(stat = "identity") + 
  facet_grid(.~patient, scales = "free", space = "free" ) +
  theme(axis.text.x = element_text(hjust = 1, vjust = 0.5, angle = 90)) + 
  scale_fill_manual(values = brewer.dark2(7))
```
```{r}
fibro3 <- merge(fibro2, subset(normal_final_srt, cells = fibrocell))
fibro3 <- harmony_rna(fibro3)
```
```{r}
DimPlot(fibro3, group.by = "cell_anno")
DimPlot(fibro3, group.by = "site")
FeaturePlot(fibro3, Fibroblast_feature)
FeaturePlot(fibro3, Endothelial_feature)
```
## T cell 
```{r}
Tcell <- subset(normal_final_srt, cell_id == "T cell")
Tcell <- harmony_rna(Tcell)
```
```{r fig.height=8, fig.width=10}
FeaturePlot(Tcell, T_cell_feature)
FeaturePlot(Tcell, Treg_feature)
FeaturePlot(Tcell, Th17_feature)
FeaturePlot(Tcell, Th1_feature)
FeaturePlot(Tcell, Naive_Th_feature)
FeaturePlot(Tcell, CTL_feature)
FeaturePlot(Tcell, CD8_effector_feature)
FeaturePlot(Tcell, CD4_T_cell_feature)
FeaturePlot(Tcell, NK_feature)
FeaturePlot(Tcell, Endothelial_feature)
FeaturePlot(Tcell, plasma_feature)
FeaturePlot(Tcell, Hepatocytes_feature)
FeaturePlot(Tcell, c("CCR7", "CD62L"))  #memory t cell
```
```{r fig.height=8, fig.width=5}
Tcell_marker <- FindAllMarkers(Tcell)
markerlist <- lapply(split(Tcell_marker, Tcell_marker$cluster), generate_genelist)
Tcell_en <- lapply(names(markerlist), function(cluster) {
  x <-markerlist[[cluster]]
  enrichment_plots(x, C8, paste(cluster, "UP"), paste(cluster, "DOWN"))
  })
names(Tcell_en) <- names(markerlist)
Tcell_en <- Tcell_en[sapply(Tcell_en, nrow)>0]
Tcell_en <- rbindlist(lapply(Tcell_en, function(x) x@result), idcol = "cluster")

```
```{r fig.height=8, fig.width=5}
Tcell$cell_anno <- "Naive T cell"
Tcell$cell_anno[Tcell$seurat_clusters == "4"] <- "Treg"
Tcell$cell_anno[Tcell$seurat_clusters == "1"] <- "CD8+ T cell"
Tcell$cell_anno[Tcell$seurat_clusters == "3"] <- "NK cell"
Idents(Tcell)<- Tcell$cell_anno

Tcell_marker <- FindAllMarkers(Tcell)
markerlist <- lapply(split(Tcell_marker, Tcell_marker$cluster), generate_genelist)
Tcell_en <- lapply(names(markerlist), function(cluster) {
  x <-markerlist[[cluster]]
  enrichment_plots(x, C8, paste(cluster, "UP"), paste(cluster, "DOWN"))
  })
names(Tcell_en) <- names(markerlist)
Tcell_en <- Tcell_en[sapply(Tcell_en, nrow)>0]
Tcell_en <- rbindlist(lapply(Tcell_en, function(x) x@result), idcol = "cluster")


```

```{r}
DimPlot(Tcell)
saveRDS(Tcell, "Tcell.Rds")
```
```{r}
df <- Tcell@meta.data %>% group_by(site, cell_anno) %>% summarise(n = n()) %>% mutate (freq = n/sum(n))

ggplot(df, aes(x = site, y = freq, fill = cell_anno)) + geom_bar(stat = "identity") + 
  theme(axis.text.x = element_text(hjust = 1, vjust = 0.5, angle = 90)) + 
  scale_fill_manual(values = brewer.dark2(7))

df <- Tcell@meta.data %>% group_by(sample, cell_anno) %>% summarise(n = n()) %>% mutate (freq = n/sum(n))
df$patient <- substr(df$sample, 1, 6 )
ggplot(df, aes(x = sample, y = freq, fill = cell_anno)) + geom_bar(stat = "identity") + 
  facet_grid(.~patient, scales = "free", space = "free" ) +
  theme(axis.text.x = element_text(hjust = 1, vjust = 0.5, angle = 90)) + 
  scale_fill_manual(values = brewer.dark2(7))
```

##B cell
```{r}
Bcell <- subset(normal_final_srt, ident = c("B cell", "Naive cell", "Plasma cell"))
Bcell <- harmony_rna(Bcell)
```
```{r}
FeaturePlot(Bcell, plasma_feature)
FeaturePlot(Bcell, T_cell_feature)
FeaturePlot(Bcell, focused_B_cell_feature)
FeaturePlot(Bcell, focused_naive_memory_B_feature)
FeaturePlot(Bcell, focused_activated_B_feature)
FeaturePlot(Bcell, Naive_Th_feature)
FeaturePlot(Bcell, NK_feature)
FeaturePlot(Bcell, c("BACH2", "EBF1", "PAX5"))
FeaturePlot(Bcell, Fibroblast_feature)
FeaturePlot(Bcell, Endothelial_feature)
FeaturePlot(Bcell, Mast_feature)
FeaturePlot(Bcell, mDC_feature)
FeaturePlot(Bcell, pDC_feature)
FeaturePlot(Bcell, Epithelial_cell)
FeaturePlot(Bcell, c("CD11B","CD14","CD33"))
```
```{r fig.height=8, fig.width=5}
Bcell_marker <- FindAllMarkers(Bcell)
markerlist <- lapply(split(Bcell_marker, Bcell_marker$cluster), generate_genelist)
Bcell_en <- lapply(names(markerlist), function(cluster) {
  x <-markerlist[[cluster]]
  enrichment_plots(x, C8, paste(cluster, "UP"), paste(cluster, "DOWN"))
  })
names(Bcell_en) <- names(markerlist)
Bcell_en <- Bcell_en[sapply(Bcell_en, nrow)>0]
Bcell_en <- rbindlist(lapply(Bcell_en, function(x) x@result), idcol = "cluster")

```

```{r}
Bcell2 <- subset(Bcell, ident = 0:7)
Bcell2 <- harmony_rna(Bcell2)
```
```{r}
FeaturePlot(Bcell2, plasma_feature)
FeaturePlot(Bcell2, T_cell_feature)
FeaturePlot(Bcell2, focused_B_cell_feature)
FeaturePlot(Bcell2, focused_naive_memory_B_feature)
FeaturePlot(Bcell2, focused_activated_B_feature)
FeaturePlot(Bcell2, Naive_Th_feature)
FeaturePlot(Bcell2, NK_feature)
FeaturePlot(Bcell2, c("BACH2", "EBF1", "PAX5"))
FeaturePlot(Bcell2, Fibroblast_feature)
FeaturePlot(Bcell2, Endothelial_feature)
FeaturePlot(Bcell2, Mast_feature)
FeaturePlot(Bcell2, mDC_feature)
FeaturePlot(Bcell2, pDC_feature)
FeaturePlot(Bcell2, Epithelial_cell)
FeaturePlot(Bcell2, c("CD11B","CD14","CD33"))

```
```{r}
Bcell2$cell_anno <- "B cell"
Bcell2$cell_anno[Bcell2$seurat_clusters == "2"] <- "Plasma cell"
Bcell2$cell_anno[Bcell2$seurat_clusters == "4"] <- "Progenitor B cell"
Idents(Bcell2) <- Bcell2$cell_anno
saveRDS(Bcell2, "Bcell.Rds")
DimPlot(Bcell2)

```

##Macrophage
```{r}
macro <- subset(normal_final_srt, ident = "Macrophage")
macro <- harmony_rna(macro)
```
```{r fig.height=10, fig.width=10}
FeaturePlot(macro, Macrophage_feature)
FeaturePlot(macro, M1_feature)
FeaturePlot(macro, M2_feature)
FeaturePlot(macro, phagocytosis_feature)
FeaturePlot(macro, angiogenesis_feature)
FeaturePlot(macro, TAM_feature)
FeaturePlot(macro, c("SPP1", "RNASE1", "MT1G", "SEPP1", "FOLR2", "NUPR1", "KLHDC8B", "CCL18", "MMP12","APOC2" ))
FeaturePlot(macro, Mast_feature)
FeaturePlot(macro, mDC_feature)
FeaturePlot(macro, Monocyte_feature)
FeaturePlot(macro, c("FCN1", "LYZ", "VCAN","CD68", "GPNMB", "CTSB")) # monocytes
FeaturePlot(macro, c("MARCO", "FABP4", "GPNMB", "CTSB"))

```

```{r fig.height=7, fig.width=5}
macro_marker <- FindAllMarkers(macro)
markerlist <- lapply(split(macro_marker, macro_marker$cluster), generate_genelist)
macro_en <- lapply(names(markerlist), function(cluster) {
  x <-markerlist[[cluster]]
  enrichment_plots(x, C8, paste(cluster, "UP"), paste(cluster, "DOWN"))
  })
names(macro_en) <- names(markerlist)
macro_en <- macro_en[sapply(macro_en, nrow)>0]
macro_en <- rbindlist(lapply(macro_en, function(x) x@result), idcol = "cluster")
```

```{r fig.height=6, fig.width=6}
DimPlot(macro, label = T)
macro$cell_anno <- "Macrophagen M0" # SAT1+
macro$cell_anno[macro$seurat_clusters == "11"] <- "Mast cell"
macro$cell_anno[macro$seurat_clusters == "9"] <- "Monocyte"
macro$cell_anno[macro$seurat_clusters %in% c("2", "10", "4")] <- "TAM"
erthcell <- WhichCells(macro, idents = "6")

FeaturePlot(macro, macro_marker %>% arrange(desc(avg_log2FC)) %>% filter(cluster == "12") %>% slice(1:6) %>% pull(gene))
table(macro$site, macro$seurat_clusters)
FeaturePlot(Bcell, c("MS4A1", "LTB", "CD37", "CD79B"))

```

```{r}
macro2 <- subset(macro, ident = c(0:5, 7:12))
macro2 <- harmony_rna(macro2)
```
```{r}
DimPlot(macro2, group.by = "cell_anno")
FeaturePlot(macro2, Macrophage_feature)
FeaturePlot(macro2, TAM_feature)
FeaturePlot(macro2, c("SPP1", "RNASE1", "MT1G", "SEPP1", "FOLR2", "NUPR1", "KLHDC8B", "CCL18", "MMP12","APOC2" ))
FeaturePlot(macro2, Mast_feature)
FeaturePlot(macro2, Monocyte_feature)
FeaturePlot(macro2, c("FCN1", "LYZ", "VCAN","CD68", "GPNMB", "CTSB")) # monocytes
FeaturePlot(macro2, c("MARCO", "FABP4", "GPNMB", "CTSB"))
```
```{r fig.height=6, fig.width=6}
macro2_marker <- FindAllMarkers(macro2)
FeaturePlot(macro2, macro2_marker %>% arrange(desc(avg_log2FC)) %>% filter(cluster == "8") %>% slice(1:6) %>% pull(gene))
FeaturePlot(macro2, macro2_marker %>% arrange(desc(avg_log2FC)) %>% filter(cluster == "11") %>% slice(1:6) %>% pull(gene))
FeaturePlot(macro2, macro2_marker %>% arrange(desc(avg_log2FC)) %>% filter(cluster == "6") %>% slice(1:6) %>% pull(gene))
FeaturePlot(macro2, macro2_marker %>% arrange(desc(avg_log2FC)) %>% filter(cluster == "5") %>% slice(1:6) %>% pull(gene))
```

```{r}
macro2$cell_anno <- "macro2phagen M0" # SAT1+
macro2$cell_anno[macro2$seurat_clusters == "9"] <- "Mast cell"
macro2$cell_anno[macro2$seurat_clusters == "6"] <- "Monocyte"
macro2$cell_anno[macro2$seurat_clusters %in% c("2")] <- "TAM"
Idents(macro2) <- macro2$cell_anno
DimPlot(macro2)

```
## Epithelial cell
```{r}
epi <- subset(normal_final_srt, ident = "Epithelial cell")
epi <- harmony_rna(epi)
```
```{r fig.height=15, fig.width=15}
load("/trigos_team/CASCADE/Analysis/signatures_analysis/objects/all_gene_sets.rda")
epi_sig <- all_gene_sets[c(1:4, 10)]
epi_sig <- lapply(epi_sig, function(x) Reduce(intersect, x))
lapply(epi_sig, function(x) FeaturePlot(epi, x))
lapply(epi_sig, function(x) VlnPlot(epi, x))
```

```{r}
FeaturePlot(epi, features = Epithelial_cell)
FeaturePlot(epi, features = c("KRT17", "KRT15", "CCL9", "CDH1"))
```

